Reduced plasticity and opportunity for selection in terrestrial ectotherm populations under climate change

### 2023-02-19 ms.rmd
Update refs because of integration with paperpile. So, cite keys have changed.

Variable thermal environments are expected to result in strong selection pressures that lead to adaptation or the evolution of phenotypic plasticity -- both of which are considered critical for population resilience to human-induced climate change [@Seebacher2015]. Phenotypic plasticity is predicted to evolve when environmental variability is high but predictable. Despite this theoretical expectation, empirical support for this prediction is scant [], likely because many organisms can behaviorally adjust micro-habitat selection to offset thermal stress, the costs of plasticity are high [], and/or the prediction is likely only supported for specific plastic responses (i.e., active and developmental plasticity). Reversible forms of phenotypic plasticity, such as acclimatization, may not be expected to adhere to such predictions.

Periods of past climatic change have had disproportionate impacts on some ecosystems over others leading to debates over which ecosystems will be most vulnerable to contemporary climate change. Studies have highlighted species occupying terrestrial ecosystems as being particularly vulnerable given their weak acclimation abilities and greater probability of experiencing thermal extremes that overwhelm physiological homeostasis [@Seebacher2015;@Gunderson2015; @Pinsky2019]. Despite marine and freshwater ecosystems appearing to have greater physiological acclimation capacity [e.g., see @Seebacher2015], it is unclear if the magnitude of physiological adjustment is sufficient. In addition, low oxygen availability has been suggested as a major factor influencing the vulnerability of aquatic ecosystems

@@

## **Materials and Methods**
### *Literature collection*
We compiled literature on ectothermic animals that measured physiological rates (e.g., metabolic rate) at two or more temperatures after having been acclimated (or acclimatized) at these temperatures. We used data from a previous meta-analysis [@Seebacher2015] and updated @Seebacher2015's data by extracting data from suitable studies from our own searches that followed the same search protocol. More specifically, we performed a literature search on the 28th of June 2017 using the Web of Science database. We limited our search to articles or proceedings papers published in English from 2013 to 2017 [the date after @Seebacher2015 searches were conducted] using the following topic search string: *"(acclimat* AND (therm* OR temp*) NOT (plant* OR tree* OR forest* OR fung* OR mammal* OR marsup* OR bird* OR human OR exercis* OR train* OR hypoxi*))"*. We further limited to the following research areas: Anatomy Morphology; Biodiversity Conservation; Biology; Ecology; Endocrinology Metabolism; Entomology; Evolutionary Biology; Marine Freshwater Biology; Physiology; Respiratory System, Reproductive Biology, Zoology.

Our search resulted in 1,321 papers for screening in Rayyan [@ouzzani2016]. We also cross-checked papers we found in our searches with a recent paper by @Havird2020, which also updates  @Seebacher2015's dataset. We included any papers that were missed between our searches and those of @Havird2020 from the dates 2013-2017. @Havird2020 added 7 new studies between 2013-2017 (mainly because they were focused on metabolic rates), and our searches differed from theirs by only a single paper [i.e., @Bulgarella2015]. Given the physiological traits we included were broader, we had a substantial increase in additional papers that we added to @Seebacher2015's dataset. More specifically, in addition to the `r length(unique(old_papers$record_num))` papers we included from the @Seebacher2015 dataset, we extracted data from an extra `r length(unique(new_papers$record_num))` papers (with a total of `r dim(new_papers)[1]` effects) that were published between 2013 - 2017 (a `r  (length(unique(new_papers$record_num)) / length(unique(old_papers$record_num)))*100`% increase in the number of published articles). Note that @Seebacher2015 included a total of 205 publications, however, not all these contained the necessary statistics we needed to derive $Q_{10}$ effect sizes and associated sampling variances (see below). While we may have missed papers, our goal was to obtain a large representative (and unbiased) sample of acclimation research rather than a comprehensive dataset. As such, our database represents the most up-to-date dataset used by @Seebacher2015 to answer questions on acclimation across ectotherms.

@@

We extracted means, standard deviations, and sample sizes for physiological rates at the two test temperatures. If there were more than two test temperatures, we choose only the test temperatures that fell within the most likely natural range of temperatures experienced by the species in question. We extracted these data from text, tables or figures of a given paper. Data were extracted from figures using the R package *metaDigitise* [@Pick2019]. We also recorded the phylum, class, order, genus and species under study and the latitude and longitude of the population that was being studied. For studies that did not provide latitude and longitude for the population, we searched for similar studies by the lab group to identify where the population was likely to have been sourced or derived from when needed. If the population was derived from the wild, we recorded the nearest latitude and longitude of the population to the field collection site. If the animals had been sourced from a commercial supplier, we took the latitude and longitude of the supplier that the paper identified the animals to have originated from. When it was not possible to find latitude and longitude using these methods, we looked up the distribution of the species in question and took the latitude and longitude of the centroid of the species' distributional range.

@@

Following @Noble2022 we calculated a series of temperature corrected effect sizes that compared mean physiological rates ($lnRR_{Q_{10}}$) as well as the variability in physiological rates ($lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$). These effect sizes are similar to the traditional temperature coefficient ($Q_{10}$), but with formal analytical approximations for their sampling variances. Sampling variances for effect sizes allowed us to make use of traditional meta-analytic modelling approaches.

@@

To compare mean physiological rates, we calculated the log $Q_{10}$ response ratio, $lnRR_{Q_{10}}$ [@Noble2022] as follows:

@@

Where, $R_{1}$ and  $R_{2}$ are mean physiological rates and  $T_{1}$ and  $T_{2}$ are the temperatures that these rates are measured. Log transformation of this ratio makes the effect size normally distributed. Equation \@ref(eq:lnq10) is essentially a temperature corrected equivalent to the log response ratio (lnRR) [@hedges1999meta; @laj2011] when the numerator and denominator are measured at different temperatures. This allows one to compare the mean of two temperature treatments directly regardless of the temperatures that these groups have been measured. The sampling variance for equation \@ref(eq:lnq10) can be computed as follows (as described in @Noble2022):

@@

#### *Comparing variance in physiological rates*
 @nakagawa2015meta recently proposed analogous effect size estimates to *lnRR* that allow for comparisons of changes in variance between two groups, the log variance ratio (*lnVR*) and the log coefficient of variation (*lnCVR*). *lnVR* and *lnCVR* are ratios that describe the relative difference in trait variability between two groups. We refer readers to @nakagawa2015meta for the equations describing *lnVR* and *lnCVR*, but these can easily be extended to their $Q_{10}$ analogues (and associated sampling variance) as follows:

@@

Equations \@ref(eq:lnq10VR) and \@ref(eq:slnq10VR) describe the change in physiological rate variance (eqn \@ref(eq:lnq10VR)) across a 10&deg;C temperature change along with its sampling variance (eqn \@ref(eq:slnq10VR)). While this is a useful metric, as discussed by @nakagawa2015meta there is often a strong mean-variance relationship that needs to be accounted for in analysing changes in variance. As such, we calculated the coefficient of variation, which standardizes changes in variance for changes in means as follows:

@@

We recorded or derived a series of moderator variables from each study that are expected to have an impact on our effect size estimates. These included the duration of acclimation in days and acclimation type ("acclimation" or "acclimatization") given that acclimation responses are expected to depend on how long chronic temperature exposure occurs (longer exposure = better acclimation response) [@Seebacher2015]. We also recorded if the sample of animals were derived from captive or wild stocks, the life-history stage of the animals used ("adult" or "juvenile") and the habitat type ("freshwater", "marine" or "terrestrial") given that @Seebacher2015 show that these factors can impact $Q_{10}$ estimates. Physiological rate measures varied widely across the studies but could generally be grouped into discrete trait categories [@Seebacher2015]. As such, using the detailed information on the trait type, and its associated units from a given study, we categorized each effect size into one of 12 trait categories. These categories included measures of whole organism performance measures including cardiac (i.e., 'cardiac') and muscle ('muscle') function, sprint speed ('sprint') and endurance ('endurance') and metabolic rates (i.e., maximal and resting metabolic rate; max MR', 'rest MR', respectively). Studies also quantified various enzymatic reaction rates, including enzymes involved in general metabolic responses (categorized as 'metabolic enzyme'), various parts of the electron transport chain, including ATPase activity ('ATPase'), mitochondrial leak ('mito_leak') and oxidation ('mito_oxidation') as well as antioxidant enzymes ('antiox'). All other traits not falling within these categories were placed into 'other'.

@@

Temperature data was extracted using the monthly averages provided by the ERA5 climate model, available from the Copernicus climate data store [@Hersbach2020]. For each population and species in the dataset we extracted a 30-year period (1958-2022) of either surface temperature at 2 meters for both terrestrial and freshwater taxa, or sea surface temperature for the marine taxa using the *ncdf4* R package (vers. `r utils::packageVersion("ncdf4")`). We chose a 2-meter resolution because we believed that it more likely to reflects the micro-thermal environment experienced by terrestrial and freshwater ectotherms at those locations.

@@

### *Meta-Analysis*
We analysed our data using multilevel meta-analytic (MLMA) and meta-regression (MLMR) models in R (vers. `r paste0(R.Version()$major, ".", R.Version()$minor)`) using *brms* [vers. `r utils::packageVersion("brms")` @Burkner2018; @Burkner2017; @stan] and *metafor* [vers. `r utils::packageVersion("metafor")` @Viechtbauer2010].  We fit both Bayesian and frequentist approaches to ensure that our results were consistent, and to create orchard plots more easily [vers. 2.0, @Nakagawa2021c]. In addition, Bayesian methods better protect against type I errors in the presence of complex sources of non-independence [@Nakagawa2021; @Song2021]. For our Bayesian models, we ran 4 MCMC chains, each with a warm-up of 1000 followed by 4000 sampling iterations keeping every 5 iterations for a total of 3200 samples from the posterior distribution. We used flat Gaussian priors for 'fixed' effects (i.e., $N(0, 10)$) and a student t-distribution for 'random' effects (i.e., $student_{t}(3, 0, 10)$). We checked that all MCMC chains were mixing and had converged (i.e., $R_{hat} = 1$). We compared any competing models using Akaike's Information Criteria (AIC) (if frequentist) or Wantabe Information Criteria (WIC) (if Bayesian). We deemed models with the lowest IC value to be best supported if there was a $\Delta IC$ between the competing models of 2 or more. If two models were within 2 $\Delta IC$ units we went with the most parsimonious model.

#### *Multi-level Meta-analysis (MLMA) Models*
We first fit multi-level meta-analysis (MLMA) models (i.e., intercept only models) for both $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$, that included study, species, and phylogeny as random effects to account for non-independence. We also included trait as a random effect to account for trait variation within the data. Our MLMA models allowed us to partition the variation in $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ among these key sources while accounting for total sampling variance in each. This allowed us to calculate total heterogeneity [i.e., $I_{total}^2$; *sensu* @NakagawaSantos2012; @Noble2022] along with various $I^2$ metrics describing the proportion of variance explained by each random effect level [@NakagawaSantos2012]. We also present 95% prediction intervals which describe the expected distribution of effects from future studies [@Nakagawa2021c; @Noble2022].

A phylogeny was derived using the Open Tree of Life (OTL) with the *rotl* package in R [vers. `r utils::packageVersion("rotl")`, @rotl2016], and plotted using *ggtree* [vers. `r utils::packageVersion("ggtree")`, @Guangchuang2017]. We resolved all polytomies in the tree. Any missing taxa were replaced with closely related species and branch lengths were computed using Grafen's method [using power = 0.7, @Grafen1989]. We used the R packages *ape* [vers. `r utils::packageVersion("ape")`, @ape] and *phytools* [vers. `r utils::packageVersion("phytools")`, @Revell2012] to prune the tree for individual analyses and calculate phylogenetic covariance (or correlation) matrices used in meta-analytic models.

@@

After quantifying levels of heterogeneity, we fit a series of multi-level meta-regression (MLMR) models to test our key questions. In all models, we included the same random effects as we used in our MLMA models. Acclimation time varied from `r range(data_wide$acclim_period, na.rm = TRUE)[1]` to `r range(data_wide$acclim_period, na.rm = TRUE)[2]` days (mean (SD) = `r mean(data_wide$acclim_period, na.rm = TRUE)` $\pm$ `r sd(data_wide$acclim_period, na.rm = TRUE)` days), and terrestrial ectotherms were acclimated for a much shorter duration (mean (SD) = `r acclim_time[3,2]` $\pm$ `r acclim_time[3,3]`, n = `r acclim_time[3,4]`) than both freshwater (mean (SD) = `r acclim_time[1,2]` $\pm$ `r acclim_time[1,3]`, n = `r acclim_time[1,4]`) and marine species (mean (SD) = `r acclim_time[2,2]` $\pm$ `r acclim_time[2,3]`, n = `r acclim_time[2,4]`). Rates of acclimation have been shown to be faster for many terrestrial groups compared to aquatic organisms [e.g., amphibians and reptiles have faster rates of acclimation than fishes; See @Einum2023], which would make it more likely that terrestrial ectotherms would show lower post acclimation $lnRR_{Q_{10}}$. To control for these possible differences, acclimation time was mean-centered (mean = 0) and included in all our models. As such, all estimates can be interpreted as values for an average level of acclimation time (i.e., `r mean(data_wide$acclim_period, na.rm = TRUE)` days).

We first tested the degree to which acute and acclimation $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ effects varied by habitat type (i.e., terrestrial, freshwater, and marine ecosystems). Models included an interaction between effect type (i.e., acute or acclimation) and habitat. Reduced mean $lnRR_{Q_{10}}_{acclimation}$ relative to $lnRR_{Q_{10}}_{acute}$ indicates that acclimation to thermal environments results in partial compensation of physiological rates (i.e., phenotypic plasticity), whereas no differences between $lnRR_{Q_{10}}_{acute}$ and $lnRR_{Q_{10}}_{acclimation}$ suggests organisms are not capable of physiological plasticity [@Havird2020]. In contrast, a difference in $lnCVR_{Q_{10}}_{acclimation}$ relative to $lnCVR_{Q_{10}}_{acute}$ would imply that changes in between individual variation in physiological rates across 10$^{\circ}$C differ depending on whether acute or acclimation responses are measured. If the interaction between effect type and habitat was not supported, then we fit a model that only contained additive effects of effect type and habitat. Following on from these models, we subset each habitat type and explored how mean $lnRR_{Q_{10}}$ changed across traits. Within each habitat (marine, freshwater, and terrestrial) we fit a series of models that included an interaction between effect type (acute / acclimation) and trait category (as defined above). Variance in effects within trait categories appeared to vary depending on the trait type in question. Comparison of a model with and without heteroscedastic residual variance favored a model with heteroscedastic residual variance across trait categories ($\Delta AIC_{c}$; marine = 58, freshwater = 120, and terrestrial = 12). To ensure models converged we limited to trait categories for each habitat with six or more effect sizes.

@@

One explanation for why terrestrial ectotherms show minimal acclimation capacity may be related to the fact that terrestrial ectotherms, were, on average acclimated for significantly less time than ectotherms from aquatic habitats . Rates of acclimation have been shown to be faster for many terrestrial groups compared to aquatic organisms [e.g., amphibians and reptiles have higher rates of acclimation than fishes; See @Einum2023]. However, faster rates of acclimation would result in opposite patterns to those we observed -- in other words, terrestrial species would be more likely to exhibit lower $lnRR_{Q_{10}}_{acclimation}$ compared to $lnRR_{Q_{10}}_{acute}$ when controlling for acclimation time.

>>>>

Variable thermal environments are expected to result in strong selection pressures that lead to adaptation or the evolution of phenotypic plasticity -- both of which are considered critical for population resilience to human-induced climate change [@Seebacher2015-wi]. Phenotypic plasticity is predicted to evolve when environmental variability is high but predictable. Despite this theoretical expectation, empirical support for this prediction is scant [], likely because many organisms can behaviorally adjust micro-habitat selection to offset thermal stress, the costs of plasticity are high [], and/or the prediction is likely only supported for specific plastic responses (i.e., active and developmental plasticity). Reversible forms of phenotypic plasticity, such as acclimatization, may not be expected to adhere to such predictions.

Periods of past climatic change have had disproportionate impacts on some ecosystems over others leading to debates over which ecosystems will be most vulnerable to contemporary climate change. Studies have highlighted species occupying terrestrial ecosystems as being particularly vulnerable given their weak acclimation abilities and greater probability of experiencing thermal extremes that overwhelm physiological homeostasis [@Seebacher2015-wi;@Gunderson2015-hz; @Pinsky2019-fn]. Despite marine and freshwater ecosystems appearing to have greater physiological acclimation capacity [e.g., see @Seebacher2015-wi], it is unclear if the magnitude of physiological adjustment is sufficient. In addition, low oxygen availability has been suggested as a major factor influencing the vulnerability of aquatic ecosystems

@@

## **Materials and Methods**
### *Literature collection*
We compiled literature on ectothermic animals that measured physiological rates (e.g., metabolic rate) at two or more temperatures after having been acclimated (or acclimatized) at these temperatures. We used data from a previous meta-analysis [@Seebacher2015-wi] and updated @Seebacher2015-wi's data by extracting data from suitable studies from our own searches that followed the same search protocol. More specifically, we performed a literature search on the 28th of June 2017 using the Web of Science database. We limited our search to articles or proceedings papers published in English from 2013 to 2017 [the date after @Seebacher2015-wi searches were conducted] using the following topic search string: *"(acclimat* AND (therm* OR temp*) NOT (plant* OR tree* OR forest* OR fung* OR mammal* OR marsup* OR bird* OR human OR exercis* OR train* OR hypoxi*))"*. We further limited to the following research areas: Anatomy Morphology; Biodiversity Conservation; Biology; Ecology; Endocrinology Metabolism; Entomology; Evolutionary Biology; Marine Freshwater Biology; Physiology; Respiratory System, Reproductive Biology, Zoology.

Our search resulted in 1,321 papers for screening in Rayyan [@ouzzani2016-ge]. We also cross-checked papers we found in our searches with a recent paper by @Havird2020-mj, which also updates  @Seebacher2015-wi's dataset. We included any papers that were missed between our searches and those of @Havird2020-mj from the dates 2013-2017. @Havird2020-mj added 7 new studies between 2013-2017 (mainly because they were focused on metabolic rates), and our searches differed from theirs by only a single paper [i.e., @Bulgarella2015-yl]. Given the physiological traits we included were broader, we had a substantial increase in additional papers that we added to @Seebacher2015-wi's dataset. More specifically, in addition to the `r length(unique(old_papers$record_num))` papers we included from the @Seebacher2015-wi dataset, we extracted data from an extra `r length(unique(new_papers$record_num))` papers (with a total of `r dim(new_papers)[1]` effects) that were published between 2013 - 2017 (a `r  (length(unique(new_papers$record_num)) / length(unique(old_papers$record_num)))*100`% increase in the number of published articles). Note that @Seebacher2015-wi included a total of 205 publications, however, not all these contained the necessary statistics we needed to derive $Q_{10}$ effect sizes and associated sampling variances (see below). While we may have missed papers, our goal was to obtain a large representative (and unbiased) sample of acclimation research rather than a comprehensive dataset. As such, our database represents the most up-to-date dataset used by @Seebacher2015-wi to answer questions on acclimation across ectotherms.

@@

We extracted means, standard deviations, and sample sizes for physiological rates at the two test temperatures. If there were more than two test temperatures, we choose only the test temperatures that fell within the most likely natural range of temperatures experienced by the species in question. We extracted these data from text, tables or figures of a given paper. Data were extracted from figures using the R package *metaDigitise* [@Pick2019-mg]. We also recorded the phylum, class, order, genus and species under study and the latitude and longitude of the population that was being studied. For studies that did not provide latitude and longitude for the population, we searched for similar studies by the lab group to identify where the population was likely to have been sourced or derived from when needed. If the population was derived from the wild, we recorded the nearest latitude and longitude of the population to the field collection site. If the animals had been sourced from a commercial supplier, we took the latitude and longitude of the supplier that the paper identified the animals to have originated from. When it was not possible to find latitude and longitude using these methods, we looked up the distribution of the species in question and took the latitude and longitude of the centroid of the species' distributional range.

@@

Following @Noble2022-ty we calculated a series of temperature corrected effect sizes that compared mean physiological rates ($lnRR_{Q_{10}}$) as well as the variability in physiological rates ($lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$). These effect sizes are similar to the traditional temperature coefficient ($Q_{10}$), but with formal analytical approximations for their sampling variances. Sampling variances for effect sizes allowed us to make use of traditional meta-analytic modelling approaches.

@@

To compare mean physiological rates, we calculated the log $Q_{10}$ response ratio, $lnRR_{Q_{10}}$ [@Noble2022-ty] as follows:

@@

Where, $R_{1}$ and  $R_{2}$ are mean physiological rates and  $T_{1}$ and  $T_{2}$ are the temperatures that these rates are measured. Log transformation of this ratio makes the effect size normally distributed. Equation \@ref(eq:lnq10) is essentially a temperature corrected equivalent to the log response ratio (lnRR) [@hedges1999-de; @Lajeunesse2011-fq] when the numerator and denominator are measured at different temperatures. This allows one to compare the mean of two temperature treatments directly regardless of the temperatures that these groups have been measured. The sampling variance for equation \@ref(eq:lnq10) can be computed as follows (as described in @Noble2022-ty):

@@

#### *Comparing variance in physiological rates*
 @Nakagawa2015-fo recently proposed analogous effect size estimates to *lnRR* that allow for comparisons of changes in variance between two groups, the log variance ratio (*lnVR*) and the log coefficient of variation (*lnCVR*). *lnVR* and *lnCVR* are ratios that describe the relative difference in trait variability between two groups. We refer readers to @Nakagawa2015-fo for the equations describing *lnVR* and *lnCVR*, but these can easily be extended to their $Q_{10}$ analogues (and associated sampling variance) as follows:

@@

Equations \@ref(eq:lnq10VR) and \@ref(eq:slnq10VR) describe the change in physiological rate variance (eqn \@ref(eq:lnq10VR)) across a 10&deg;C temperature change along with its sampling variance (eqn \@ref(eq:slnq10VR)). While this is a useful metric, as discussed by @Nakagawa2015-fo there is often a strong mean-variance relationship that needs to be accounted for in analysing changes in variance. As such, we calculated the coefficient of variation, which standardizes changes in variance for changes in means as follows:

@@

We recorded or derived a series of moderator variables from each study that are expected to have an impact on our effect size estimates. These included the duration of acclimation in days and acclimation type ("acclimation" or "acclimatization") given that acclimation responses are expected to depend on how long chronic temperature exposure occurs (longer exposure = better acclimation response) [@Seebacher2015-xj]. We also recorded if the sample of animals were derived from captive or wild stocks, the life-history stage of the animals used ("adult" or "juvenile") and the habitat type ("freshwater", "marine" or "terrestrial") given that @Seebacher2015-xj show that these factors can impact $Q_{10}$ estimates. Physiological rate measures varied widely across the studies but could generally be grouped into discrete trait categories [@Seebacher2015-xj]. As such, using the detailed information on the trait type, and its associated units from a given study, we categorized each effect size into one of 12 trait categories. These categories included measures of whole organism performance measures including cardiac (i.e., 'cardiac') and muscle ('muscle') function, sprint speed ('sprint') and endurance ('endurance') and metabolic rates (i.e., maximal and resting metabolic rate; max MR', 'rest MR', respectively). Studies also quantified various enzymatic reaction rates, including enzymes involved in general metabolic responses (categorized as 'metabolic enzyme'), various parts of the electron transport chain, including ATPase activity ('ATPase'), mitochondrial leak ('mito_leak') and oxidation ('mito_oxidation') as well as antioxidant enzymes ('antiox'). All other traits not falling within these categories were placed into 'other'.

@@

Temperature data was extracted using the monthly averages provided by the ERA5 climate model, available from the Copernicus climate data store [@Hersbach2020-hx]. For each population and species in the dataset we extracted a 30-year period (1958-2022) of either surface temperature at 2 meters for both terrestrial and freshwater taxa, or sea surface temperature for the marine taxa using the *ncdf4* R package (vers. `r utils::packageVersion("ncdf4")`). We chose a 2-meter resolution because we believed that it more likely to reflects the micro-thermal environment experienced by terrestrial and freshwater ectotherms at those locations.

@@

### *Meta-Analysis*
We analysed our data using multilevel meta-analytic (MLMA) and meta-regression (MLMR) models in R (vers. `r paste0(R.Version()$major, ".", R.Version()$minor)`) using *brms* [vers. `r utils::packageVersion("brms")` @Burkner2018-ur; @Burkner2017-wg; @noauthor_2021-az] and *metafor* [vers. `r utils::packageVersion("metafor")` @Viechtbauer2010-fn].  We fit both Bayesian and frequentist approaches to ensure that our results were consistent, and to create orchard plots more easily [vers. 2.0, @Nakagawa2021-ls]. In addition, Bayesian methods better protect against type I errors in the presence of complex sources of non-independence [@Nakagawa2021-uf; @Song2021-ba]. For our Bayesian models, we ran 4 MCMC chains, each with a warm-up of 1000 followed by 4000 sampling iterations keeping every 5 iterations for a total of 3200 samples from the posterior distribution. We used flat Gaussian priors for 'fixed' effects (i.e., $N(0, 10)$) and a student t-distribution for 'random' effects (i.e., $student_{t}(3, 0, 10)$). We checked that all MCMC chains were mixing and had converged (i.e., $R_{hat} = 1$). We compared any competing models using Akaike's Information Criteria (AIC) (if frequentist) or Wantabe Information Criteria (WIC) (if Bayesian). We deemed models with the lowest IC value to be best supported if there was a $\Delta IC$ between the competing models of 2 or more. If two models were within 2 $\Delta IC$ units we went with the most parsimonious model.

#### *Multi-level Meta-analysis (MLMA) Models*
We first fit multi-level meta-analysis (MLMA) models (i.e., intercept only models) for both $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$, that included study, species, and phylogeny as random effects to account for non-independence. We also included trait as a random effect to account for trait variation within the data. Our MLMA models allowed us to partition the variation in $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ among these key sources while accounting for total sampling variance in each. This allowed us to calculate total heterogeneity [i.e., $I_{total}^2$; *sensu* @Nakagawa2012-su; @Noble2022-ty] along with various $I^2$ metrics describing the proportion of variance explained by each random effect level [@Nakagawa2012-su]. We also present 95% prediction intervals which describe the expected distribution of effects from future studies [@Nakagawa2021-ls; @Noble2022-ty].

A phylogeny was derived using the Open Tree of Life (OTL) with the *rotl* package in R [vers. `r utils::packageVersion("rotl")`, @Michonneau2016-if], and plotted using *ggtree* [vers. `r utils::packageVersion("ggtree")`, @Yu2017-wj]. We resolved all polytomies in the tree. Any missing taxa were replaced with closely related species and branch lengths were computed using Grafen's method [using power = 0.7, @Grafen1989-pr]. We used the R packages *ape* [vers. `r utils::packageVersion("ape")`, @Paradis2019-gx] and *phytools* [vers. `r utils::packageVersion("phytools")`, @Revell2012-vq] to prune the tree for individual analyses and calculate phylogenetic covariance (or correlation) matrices used in meta-analytic models.

@@

After quantifying levels of heterogeneity, we fit a series of multi-level meta-regression (MLMR) models to test our key questions. In all models, we included the same random effects as we used in our MLMA models. Acclimation time varied from `r range(data_wide$acclim_period, na.rm = TRUE)[1]` to `r range(data_wide$acclim_period, na.rm = TRUE)[2]` days (mean (SD) = `r mean(data_wide$acclim_period, na.rm = TRUE)` $\pm$ `r sd(data_wide$acclim_period, na.rm = TRUE)` days), and terrestrial ectotherms were acclimated for a much shorter duration (mean (SD) = `r acclim_time[3,2]` $\pm$ `r acclim_time[3,3]`, n = `r acclim_time[3,4]`) than both freshwater (mean (SD) = `r acclim_time[1,2]` $\pm$ `r acclim_time[1,3]`, n = `r acclim_time[1,4]`) and marine species (mean (SD) = `r acclim_time[2,2]` $\pm$ `r acclim_time[2,3]`, n = `r acclim_time[2,4]`). Rates of acclimation have been shown to be faster for many terrestrial groups compared to aquatic organisms [e.g., amphibians and reptiles have faster rates of acclimation than fishes; See @Einum2023-lz], which would make it more likely that terrestrial ectotherms would show lower post acclimation $lnRR_{Q_{10}}$. To control for these possible differences, acclimation time was mean-centered (mean = 0) and included in all our models. As such, all estimates can be interpreted as values for an average level of acclimation time (i.e., `r mean(data_wide$acclim_period, na.rm = TRUE)` days).

We first tested the degree to which acute and acclimation $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ effects varied by habitat type (i.e., terrestrial, freshwater, and marine ecosystems). Models included an interaction between effect type (i.e., acute or acclimation) and habitat. Reduced mean $lnRR_{Q_{10}}_{acclimation}$ relative to $lnRR_{Q_{10}}_{acute}$ indicates that acclimation to thermal environments results in partial compensation of physiological rates (i.e., phenotypic plasticity), whereas no differences between $lnRR_{Q_{10}}_{acute}$ and $lnRR_{Q_{10}}_{acclimation}$ suggests organisms are not capable of physiological plasticity [@Havird2020-mj]. In contrast, a difference in $lnCVR_{Q_{10}}_{acclimation}$ relative to $lnCVR_{Q_{10}}_{acute}$ would imply that changes in between individual variation in physiological rates across 10$^{\circ}$C differ depending on whether acute or acclimation responses are measured. If the interaction between effect type and habitat was not supported, then we fit a model that only contained additive effects of effect type and habitat. Following on from these models, we subset each habitat type and explored how mean $lnRR_{Q_{10}}$ changed across traits. Within each habitat (marine, freshwater, and terrestrial) we fit a series of models that included an interaction between effect type (acute / acclimation) and trait category (as defined above). Variance in effects within trait categories appeared to vary depending on the trait type in question. Comparison of a model with and without heteroscedastic residual variance favored a model with heteroscedastic residual variance across trait categories ($\Delta AIC_{c}$; marine = 58, freshwater = 120, and terrestrial = 12). To ensure models converged we limited to trait categories for each habitat with six or more effect sizes.

@@

One explanation for why terrestrial ectotherms show minimal acclimation capacity may be related to the fact that terrestrial ectotherms, were, on average acclimated for significantly less time than ectotherms from aquatic habitats . Rates of acclimation have been shown to be faster for many terrestrial groups compared to aquatic organisms [e.g., amphibians and reptiles have higher rates of acclimation than fishes; See @Einum2023-lz]. However, faster rates of acclimation would result in opposite patterns to those we observed -- in other words, terrestrial species would be more likely to exhibit lower $lnRR_{Q_{10}}_{acclimation}$ compared to $lnRR_{Q_{10}}_{acute}$ when controlling for acclimation time.
